RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-31 casestudies.tex
yes

% IMPORTANT: what is unique about RCloud here

% From prototyping to dashboard

% Getting other information from the web

%

>>>>

We present two example applications of the capabilities provided by RCloud
based on synthetic data and applications.\footnote{Unfortunately, we could
not show production notebooks because of concerns about revealing proprietary
or confidential information.} The main pont of the examples is that they are
representative of a progression through the visual analytics development
and deployment process that we aim to support.

\subsection{Stock Price Analysis\label{sec:stockvis}}

\begin{figure}
\includegraphics[width=\linewidth]{fig/casestudy1/casestudy1.pdf}
\caption{\label{sec:stockvis}Iterations of a stock ticker
  visualization, after an example by Hadley Wickham.  A
  simple, static visualization of the closing price of a single ticker
  is progressively developed into a configurable display suitable
  for dashboarding, then into an interactive visualization of the
  volatility and volume of two stocks, then finally into an API
  for data access by other RCloud notebooks. All notebooks
  in this example can be loaded as web pages. When a notebook
  corresponding to a function call is displayed a web page,
  its associated documentation is displayed.}
\end{figure}

Our first example is a sequence of visualizations of the performance
of financial stocks over a multi-year period. The first visualization
in this example uses ggplot2 and is due to Hadley Wickham. It reads
data provided by Yahoo! Finance as a web service, and shows the price
for a single trading symbol.

The webpage that produces that visualization includes a link back to
the generating source code, from which a user can fork the notebook
and, in this case, add a configurable ticker based on the URL of
the notebook. Notice how the notebook changes very little.

From there, the notebook author (or another user) may decide to
provide convenient access to the data apart from its visualization.
This is achieved by creating a notebook that defines a function.
This notebook then becomes a \emph{subroutine} for other notebooks,
and is version-controlled like other notebooks.

\todo{Interactive version.}

%% This notebook can then be called in an \emph{interactive}
%% visualization, which uses the Javascript

%% Easy to convert it into configurable entry point for dashboarding.

%% Now consider an analyst that wants to understand the price dynamics of
%% different stock attributes, different points in time, etc. For that,
%% interactive visualization is very attractive. describe dcplot and
%% notebook which uses it.

%% Then describe abstraction of stock-data fetching, to talk about
%% calling notebooks from other notebooks.

%% then call.r?

\subsection{text analysis\ref{sec:textvis}}

Same.

% IMPORTANT: what is unique about RCloud here
% From prototyping to dashboard
% Getting other information from the web
%

### 2014-03-31 conclusion.tex
yes

We presented the ideas that motivated the design of RCloud,
a prototype visual analytics environment to support
collaboration, searching, sharing, and publishing of
experiments, similar to DevOps development of web services.
Experience with the prototype provided evidence that
data science teams and the organizations in which they work
can benefit from having such features.

>>>>

We described the motivation introducing web technologies
for searching, collaborating, recommending, publishing
reusable notebooks as part of the visual analytics process.
We implemented these ideas in the RCloud prototype.
Rcloud assists software development and release using
an approach similar to DevOps in web services.
Experience with the prototype provided evidence that
data science teams, and the organizations in which they work,
benefit from having such capabilities.

### 2014-03-31 system.tex
yes

Edit

View/Run/Share

Call

Browse/Search. ``More like this'', and a custom webpage solely for browsing.

@@

The basic unit of operation in RCloud is a \emph{notebook}. RCloud
notebooks hold a sequence of \emph{cells}, each of which represents a
snippet of code or hypertext in Markdown. This is not a novel idea;
executable documents structured in this way are present in a wide
variety of present software systems, from Mathematica to IPython to
Sage and others.

One of the main differences between RCloud and other systems is the
notion that notebooks are ``always deployed''. This has a variety of
advantages which we discuss \todo{where?}. On the other hand, there
are a variety of situations that might ask for a specific version of a
notebook. Rather than have the notebook designer decide ahead of time
which versions should be preserved, we choose to expose
\emph{transparent} versioning. This is similar to models like
Jankun-Kelly et al.'s p-set calculus \todo{cite} and VisTrails's
version tree \todo{cite}, where every change in the state of the
system is tracked.

Implementing the entire version-control subsystem would be a large
undertaking. Instead, we build the infrastructure on top of Github's
\emph{gists}~\todo{cite}. Github offers a HTTP interface
for creating simplified git repositories (the main limitation is the
use of text-only files and a single directory). The GitHub web-service
API provides much of the infrastructure we need for the versioning
portion of the storage back end: access to previous versions,
comments, starring, and forking.

Using GitHub was mainly a tactical decision to reduce our
time-to-prototype. At the same time, it had the fortuitous decision of
forcing our backend to communicate with separate processes over HTTP
and JSON. This had the serendipitous advantage of quickly enabling
full search over RCloud notebooks (built using Apache SOLR).
%
This perspective has implications for
systems-based research and prototyping in visual analytics.
%
In order for small teams to efficiently explore ideas in the present
software ecosystem, it pays off to use as many existing solutions as
possible.
%
Even though it involved learning a variety of standards that are not
directly related to visual analytics, our early decision to use the
GitHub gist API has been quite fruitful. We expect similar payoffs to
happen for related projects as well.

@@

We incorporate both explicit and implicit indications of interest
in notebooks. Explicit interest is signaled by ``starring,'' or
clicking on a button that marks a notebook as interesting.
This makes explicit indication of interest a nearly trivial operation,
always readily available, thus encouraging its use.

@@

HTTP and web services are convenient but much of the protocol is
stateless. (For example, GET requests are required to not change the
remote state, and results can be cached in transparent proxies along
the network). In the case of close communication between a running
R process and a web browser, we seek something that is more flexible
than distinct URLs for every R process (with the resulting
complication in providing access

>>>>

\todo{Edit}

\todo{View/Run/Share}

\todo{Call}

\todor{Browse/Search}. ``More like this'', and a custom webpage solely for browsing.

@@

The basic unit of operation in RCloud is a \emph{notebook}. A notebook
holds a sequence of \emph{cells}, each of which represents a
snippet of code or hypertext in Markdown. This is not a novel idea;
executable documents structured in this way are present in a wide
variety of other software systems, including Mathematica, IPython and Sage.

One of the main differences between RCloud and other systems is the
notion that notebooks are ``always deployed''. This has a variety of
advantages which we discuss \todo{where?}. On the other hand, there
are a variety of situations that might ask for a specific version of a
notebook. Rather than have the notebook designer decide ahead of time
which versions should be preserved, we embrace \emph{transparent} versioning.
This is similar to models like Jankun-Kelly et al.'s p-set calculus \todo{cite}
and VisTrails's version tree \todo{cite}, where every change in the state of the
system is tracked.

Implementing the entire version-control subsystem would be a large
undertaking. Fortunately, we can build RCloud on top of Github's
\emph{gists}~\todo{cite}. Github offers a HTTP interface
for creating simplified git repositories (the main limitation being the
use of text-only files in a single directory). The GitHub web-service
API provides most of the semantics we need for the versioning
portion of the storage back end: access to previous versions,
comments, starring, and forking.

Using GitHub was mainly a tactical decision to reduce our
time-to-prototype. At the same time, it had the beneficial effect of
forcing our backend to communicate with separate processes over HTTP
and JSON. This had the serendipitous advantage of quickly enabling
full search over RCloud notebooks (using Apache SOLR).
%
This perspective has implications for
systems-based research and prototyping in visual analytics.
%
In order for small teams to efficiently explore ideas in the present
software ecosystem, it pays off to use as many existing solutions as
possible.
%
Even though it involved learning a variety of standards that are not
directly related to visual analytics, our early decision to use the
GitHub gist API has been quite fruitful. We expect related projects
can gain similar benefits by adopting the same strategy.

@@

We incorporate both explicit and implicit indications of interest
in notebooks. Explicit interest is signaled by ``starring,'' or
clicking on a button that marks a notebook as interesting.
This makes explicit indication of interest a nearly trivial operation,
always readily available, to encourage its use.

@@

HTTP and web services are convenient but much of the protocol is
stateless. (For example, GET requests are required to not change the
remote state, and results can be cached in transparent proxies along
the network). In the case of close communication between a running
R process and a web browser, we need something that is more flexible
than distinct URLs for every R process (with the resulting
complication in providing access
